\documentclass{article}

\usepackage{graphviz}
\usepackage{fullpage}
\usepackage{parskip}
\usepackage{url}
\usepackage{hyperref}

\usepackage{listings}
\lstset{
	numbers=left,
	basicstyle=\footnotesize,
	captionpos=b,
	xleftmargin=0.3in
}

\usepackage[
	backend=biber,
	autocite=footnote,
	bibstyle=authortitle,
	citestyle=verbose-inote,
]{biblatex}
\addbibresource{main.bib}
\setlength\bibitemsep{1em}

\renewcommand{\labelitemi}{-}

\begin{document}

\title{Interview Questions}
\author{
	\Large{Jeremiah Mahler}\\
	\small{\href{mailto:jmmahler@gmail.com}{\textless jmmahler@gmail.com\textgreater}}
}
\date{\today}
\maketitle
%\clearpage

%\thispagestyle{empty}
%\tableofcontents
%\clearpage

\section{Introduction}

This is a collection of questions that I have been asked, or have asked,
during job interviews.
The technical questions are related to Electrical Engineering and Programming.
And included are notes describing how I would answer a question as
well as what I am looking for in an answer.

\section{Questions For Managers}

In general I am looking for a manager who leads has employees and gives
them the freedom to do their best work.

\begin{enumerate}

\item If I was hired for this position what would be the first project that
I would work on?

\item What are your favorite books that have influenced your management style?
	\begin{itemize}
	\item Peopleware by Tom DeMarco
	\item Getting Things Done by David Allen
	\item Drive by Daniel Pink
	\item The Mythical Man Month by Frederick P. Brooks
	\end{itemize}

\item Tell me about a time when there was a disagreement among your employees.
	\begin{itemize}
	\item Was he able to convince them without asserting his authority?
	\end{itemize}

\item Tell me about a time when one of your employees disagreed with you.
	\begin{itemize}
	\item Was he able to convince them without asserting his authority?
	\end{itemize}

\item (similar) Have you ever had to tell an employee to do something that
they disagreed with you about?

\item If an employee did something different than what you told them to do,
	how would you respond?
	\begin{itemize}
	\item Does he trust his employees?
	\item Or is he threatened by a loss of control?
	\end{itemize}

\item (follow up) What if the solution they came up with was significantly
	better?

\item If you had no authority over your employees would they still do
	what you ask?
	\begin{itemize}
	\item A boss or a leader?
	\item A boss leverages his authority to force employees to do what he wants.
	\item People will follow a leader even if they have no authority.
	\end{itemize}

\end{enumerate}

\section{Questions From Managers}

\begin{enumerate}

\item Where do you see yourself in 5 years?
	\begin{itemize}
	\item Should have a specific goal that they are trying to achieve.
		And the position being sought should aligh with that goal.
	\end{itemize}

\item Do you have any questions for me?
	\begin{itemize}
	\item Should have lots of questions which shows an interest in the
		company and the position.  (see "Questions For Managers"
		section)
	\end{itemize}

\item Why would you be a good candidate?
	\begin{itemize}
	\item There are many reasons why I would be a good candidate.
	But there are probably only a few that you are interested in.
	Describe to me your ideal candidate and I will tell you if I have
	anything in common.
	\end{itemize}

\end{enumerate}

\section{Questions From Electrical Engineers and Programmers}

\begin{enumerate}

\item What is your favorite programming language?
	\begin{itemize}
	\item My favorite language is Lisp because it changed the way I
		think about programming.  Things such as: Prefix notation,
		anonymous functions, tail recursion, functional programming,
		macros, linked lists, etc.
	\end{itemize}

\item What is a memory leak?
	\begin{itemize}
	\item A memory leak occurs when memory is allocated but is not
		freed.  If this leak occurs inside a loop it can quickly consume
		all the memory in the system.

		Most often this occurs in C when a malloc is performed without a
		corresponding free.  There are also functions which internally
		allocate memory and return a pointer.  These must be freed to
		avoid a leak.

  		A good tool for finding memory leaks when using Gcc is Valgrind.
	\end{itemize}

\item Tell me about a challenging problem that kept you up at night.

\item Why did you choose Electrical Engineering?
	\begin{itemize}
	\item I considered several degrees which were of interest to me:
		Computer Science, Computer Engineering and Electrical
		Engineering.  I already had a lot of experience programming so
		Computer Science would have been the least challenging.
		However, the low level circuits and electronics were interesting
		to me as well.  I chose EE because it was the most challenging
		degree that was of interest to me.
	\end{itemize}

\item Plot the magnitude of a 16-bit signed and unsigned number.
	\begin{itemize}
	\item \href{http://jmahler.github.io/programming/2014/01/17/sign_magnitude.html}{Magnitude of Signed/Unsigned 16-bit}
	\end{itemize}

\end{enumerate}

\section{Questions For Electrical Engineers and Programmers}

\begin{enumerate}

\item Have you ever written good quality C code that used a goto statement?
	\begin{itemize}
	\item The general rule is to never use goto statements.  But in
		Linux device drivers they are used to undo allocated resources
		during module insertion.  This is the one place where a goto is
		acceptable.
	\end{itemize}

\item If your manager had no authority to fire you would you still do
	what he asks?
	\begin{itemize}
	\item A boss or a leader?
	\end{itemize}

\item Tell me about a time when you disagreed with your manager.
	\begin{itemize}
	\item Did his manager assert his authority to get him to do something he
		did not agree with?
	\end{itemize}

\end{enumerate}

\section{Questions For Others, General}

\begin{enumerate}

\item What is a typical day for you like?

\item What surprised you the most about working here that you did
not anticipate before you were hired?

\item Are you allowed any time to work on your own ideas?
	\begin{itemize}
	\item Google 20\% time.  A great way to keep employees enthusiastic
		about the projects they are working on.  And the company often
		benefits from their good ideas.
	\end{itemize}

\item Tell me about some of the projects you have worked on here.

\end{enumerate}

\section{Questions From Recruiters}

\begin{enumerate}

\item Why do you want to work for this company?  Why not competitor X or Y?
	\begin{itemize}
	\item (Tesla) The benefits of the electric car on our society is
		obvious (cleaner air, reduced dependence on foreign oil,  more
		efficient, etc).  And yet most manufacturers choose not to
		develop an electric car even though they have more than enough
		resources to do so.  I don't want to work for any manufacturer
		where profit is the primary goal and the long term health of our
		society is secondary.
	\end{itemize}

\item Why do you think you would be a good candidate at this company?

\end{enumerate}

\clearpage
\printbibliography[heading=bibintoc]

\end{document}
